---
title: "Master"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
library(GGally)
library(caret) # for cross validation
library(readr)
library(dataMaid)
library(corrplot)
```

```{r "data import", include=FALSE}

load("C:/Users/rasmu/OneDrive - Aalborg Universitet/0P5/R-projekt-p5/HOME dataset updated/HOME_med_koordinater.Rda")
#Tænker at man kan nøjes med at inkludere kode og plots ved brug af \includepdf i overleaf, altså ved at slette den smule lim tekst jeg skriver mellem code chunks og så knitte til pdf. Eller bare kopiere relevante chunks ind i andet dokument for at knitte.
```

Firstly dataMaid is used to get a generel feel for the data, this will not be included in the project as it is too large.

```{r "data report dataMaid", eval=FALSE, include=FALSE}
# makeDataReport(HOME)
```

From the data report a number of issues are noticed with the classes of the parameters, such as parameters which should be integers and so on. Aswell as parameters with a lot of missing values.

```{r "fixing parameter classes"}
#select only "ejerlejligheder" in "Aalborg" and remove parameters with too many NA's, and removing rows with NA's

EL_HOME <- HOME %>% filter(EjdType == "Ejerlejlighed") %>% 
  filter(Postnr == 9000) %>% select(-c(NytKokken, NyeVinduer, NytBadevaerelse, AntalSovevaerelser, TaetVedVand)) %>% na.omit()


EL_HOME$OmbygningSket <- factor(EL_HOME$OmbygningSket)
EL_HOME$Hoejhus <- factor(EL_HOME$Hoejhus)
EL_HOME$Storgrund <- factor(EL_HOME$Storgrund)
EL_HOME$Kaelder <- factor(EL_HOME$Kaelder)
EL_HOME$MangeToil <- factor(EL_HOME$MangeToil)
EL_HOME$Salgsaar <- as.integer(EL_HOME$Salgsaar)
EL_HOME$GarageCarport <- factor(EL_HOME$GarageCarport)
EL_HOME$Altan <- factor(EL_HOME$Altan)
EL_HOME$AntalToiletter <- factor(EL_HOME$AntalToiletter)
EL_HOME$AntalPlan <- factor(EL_HOME$AntalPlan)
EL_HOME$Boligtilstand <- factor(EL_HOME$Boligtilstand)
EL_HOME$EjdType <- factor(EL_HOME$EjdType)
EL_HOME$Liggetid <- as.integer(EL_HOME$Liggetid)


```

Now the distribution of "Kontant pris" is checked, which should be normal for a GLM

```{r "check normal distribution"}

qqnorm(EL_HOME$Kontantpris)
qqline(EL_HOME$Kontantpris)

# A logarithmic transformation is tried
qqnorm(log(EL_HOME$Kontantpris))
qqline(log(EL_HOME$Kontantpris))

ggplot(EL_HOME, aes(log(Kontantpris))) +
  geom_density(size = 1.2)

ggplot(EL_HOME, aes((Kontantpris))) +
  geom_density(size = 1.2)

```

Parameter for season.

```{r "parameters for season and floor"}
# Gathering quarter parameters to one
EL_HOME$Season <- EL_HOME$Kvartal1 + EL_HOME$Kvartal2 * 2 + EL_HOME$Kvartal3 * 3 + 
  EL_HOME$Kvartal4 *4
EL_HOME$Season <- factor(EL_HOME$Season)

```

Firstly a model with all relevant parameters is tried

```{r "initial models"}


mod1 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Opfoerelsesaar + Liggetid + Salgsaar + Alder + Season, data = EL_HOME)
summary(mod1)

# attempt at removal of insignificant parameters
# season, liggetid
mod2 <- lm(log(Kontantpris) ~ Boligtilstand + Boligareal + Altan +
             OmbygningSket + Opfoerelsesaar + Salgsaar + Alder, data = EL_HOME)
summary(mod2)
anova(mod1,mod2)
# the p-value is above 0.05 so the two parameters can safely be removed

```

It seems from the summary of mod2 that no other paramters are insignificant.

```{r "residual plots for chosen model"}

res <- rstudent(mod2)
pred <- predict(mod2)

ggplot(cbind(EL_HOME, res, pred), aes(Kontantpris, res)) +
  geom_point() +
  theme_bw()

ggplot(cbind(EL_HOME, res, pred), aes(pred, res)) +
  geom_point() +
  theme_bw()

hat <- data.frame(hatvalues(mod2))

ggplot(hat, aes(hat$hatvalues.mod2.)) +
  geom_histogram() +
  theme_bw()

hat_Large <- which(hat$hatvalues.mod2.> 0.1)

EL_HOME <- EL_HOME[-hat_Large,] # after this repeat chunk "initial models" until no more outliers

summary(EL_HOME$Boligtilstand)
EL_HOME$Boligtilstand <- factor(EL_HOME$Boligtilstand)
```


Cross validation K fold

```{r "cross validation k fold"}

train.control <- trainControl(method = "repeatedcv", 
                              number = 5, repeats = 10)
# Train the model
model <- train(log(Kontantpris) ~ Boligtilstand + Boligareal + Altan +
                 OmbygningSket + Opfoerelsesaar + Salgsaar + Alder, data = EL_HOME, method = "lm",
               trControl = train.control)
# Summarize the results
print(model)


model$finalModel
mod2$coefficients
model$results

```

cross validation leave on out

```{r }




```









