---
title: "Data analysis"
always_allow_html: true
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---
\pagenumbering{gobble}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
library(GGally)
library(caret) # for cross validation
library(readr)
library(dataMaid)
library(corrplot)
library(leaflet)
library(grid)
```

```{r "data import", include=FALSE}
load("C:/Users/rasmu/OneDrive - Aalborg Universitet/0P5/R-projekt-p5/HOME dataset updated/HOME_med_koordinater.Rda")
#Tænker at man kan nøjes med at inkludere kode og plots ved brug af \includepdf i overleaf, altså ved at slette den smule lim tekst jeg skriver mellem code chunks og så knitte til pdf. Eller bare kopiere relevante chunks ind i andet dokument for at knitte.
```


Firstly dataMaid is used to get a generel feel for the data, this will not be included in the project as it is too large.

```{r "data report dataMaid", eval=FALSE, include=FALSE}
# makeDataReport(HOME)
```

```{r}

FULL_HOME <- HOME %>% filter(EjdType == "Ejerlejlighed") %>% 
  select(-c(NytKokken, NyeVinduer, NytBadevaerelse, 
            AntalSovevaerelser, TaetVedVand)) %>% na.omit()



```


From the data report a number of issues are noticed with the classes of the parameters, such as parameters which should be integers and so on. Aswell as parameters with a lot of missing values.

```{r "fixing parameter classes1"}
# select only "ejerlejligheder" in "Aalborg" and
# remove parameters with too many NA's, and removing rows with NA's

EL_HOME <- HOME %>% filter(EjdType == "Ejerlejlighed") %>% 
  filter(Postnr == 9000) %>% 
  select(-c(NytKokken, NyeVinduer, NytBadevaerelse, 
            AntalSovevaerelser, TaetVedVand)) %>% na.omit()
```
\newpage

```{r "fixing parameter classes2"}
pal <- colorQuantile("YlOrRd", NULL, n = 5)
leaflet(EL_HOME) %>% 
  addTiles() %>%
  addCircleMarkers(color = ~pal(Kontantpris), radius = 1)
```
\newpage

```{r "fixing parameter classes3"}
EL_HOME$OmbygningSket <- factor(EL_HOME$OmbygningSket)
EL_HOME$Hoejhus <- factor(EL_HOME$Hoejhus)
EL_HOME$Storgrund <- factor(EL_HOME$Storgrund)
EL_HOME$Kaelder <- factor(EL_HOME$Kaelder)
EL_HOME$MangeToil <- factor(EL_HOME$MangeToil)
EL_HOME$Salgsaar <- as.integer(EL_HOME$Salgsaar)
EL_HOME$GarageCarport <- factor(EL_HOME$GarageCarport)
EL_HOME$Altan <- factor(EL_HOME$Altan)
EL_HOME$AntalToiletter <- factor(EL_HOME$AntalToiletter)
EL_HOME$AntalPlan <- factor(EL_HOME$AntalPlan)
EL_HOME$Boligtilstand <- factor(EL_HOME$Boligtilstand)
EL_HOME$EjdType <- factor(EL_HOME$EjdType)
EL_HOME$Liggetid <- as.integer(EL_HOME$Liggetid)


```

Now the distribution of "Kontant pris" is checked, which should be normal for a GLM.

```{r "check normal distribution1"}
ggplot(EL_HOME, aes(sample = Kontantpris)) +
  stat_qq() +
  stat_qq_line()
```
\newpage

```{r "check normal distribution3"}
# A logarithmic transformation is tried
ggplot(EL_HOME, aes(sample = log(Kontantpris))) +
  stat_qq() +
  stat_qq_line()
```
\newpage

```{r "check normal distribution5"}
ggplot(EL_HOME, aes(log(Kontantpris))) +
  geom_density(size = 1.2) +
  theme_bw()
```
\newpage

```{r "check normal distribution6"}
ggplot(EL_HOME, aes((Kontantpris))) +
  geom_density(size = 1.2) +
  theme_bw()
```
\newpage

# Parameter for season.

Also removing parameters irrelevant for condoes

```{r "parameter for season"}
# Gathering quarter parameters to one
EL_HOME$Season <- EL_HOME$Kvartal1 + EL_HOME$Kvartal2 * 2 + EL_HOME$Kvartal3 * 3 + 
  EL_HOME$Kvartal4 *4
EL_HOME$Season <- factor(EL_HOME$Season)


centrum <- c(9.921747, 57.048820)
data_m_centrum <- data.frame(long = EL_HOME$long - centrum[1], lat = EL_HOME$lat
                             - centrum[2])
length_centrum <- apply(data_m_centrum, 1, norm, type = "2")
EL_HOME$Fracentrum <- length_centrum

# removing Kvartal parameters, and also other irrelevant parameters for our subset
# of the data
EL_HOME <- EL_HOME %>% select(-c(Kvartal1, Kvartal2, Kvartal3, Kvartal4,
                                 GarageCarport, x, y, long, lat, Hoejhus,
                                 Storgrund, Kaelder, Kaelderareal, Grundareal,
                                 AntalPlan))

```

# Generel Linear model

Firstly a model with all relevant parameters is tried. For example Alder is removed because it isn't linearly independent of Opfoerelsesaar and Salgsaar.

```{r "initial model1"}
mod1 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Opfoerelsesaar + Liggetid + Salgsaar + Season +
             Fracentrum,
           data = EL_HOME)
```
\newpage

```{r "initial model2"}
summary(mod1)
```
\newpage
We will now check for outliers as a result of mistakes in the data.


```{r "residuals and leverage1", warning=FALSE}
res <- rstudent(mod1)
pred <- predict(mod1)
index <- seq_along(EL_HOME$Kontantpris)
ggplot(cbind(EL_HOME, res, pred, index), aes(index, res)) +
  geom_point() +
  theme_bw()
```
\newpage


```{r "residuals and leverage2", warning=FALSE}
ggplot(cbind(EL_HOME, res, pred), aes(pred, res)) +
  geom_point() +
  theme_bw()
```
\newpage

```{r "residuals and leverage3", warning=FALSE}
res_large <- which(abs(res) > 4.5)
EL_HOME[res_large,]

hat <- data.frame(hatvalues(mod1))

ggplot(hat, aes(hat$hatvalues.mod1.)) +
  geom_histogram() +
  theme_bw()
```
\newpage

```{r "residuals and leverage4", warning=FALSE}
hat_large <- which(hat > 0.2)
EL_HOME[hat_large,]
#these condoes seem to be mistakes so are removed
EL_HOME <- EL_HOME[-res_large,]
EL_HOME <- EL_HOME[-hat_large,]
# the model is then reduced
```
\newpage

```{r "residuals and leverage5", warning=FALSE}
# Some factor parameters might have changed
summary(EL_HOME$Boligtilstand)
```
\newpage

```{r "residuals and leverage6", warning=FALSE}
# Factor with only 1 entry is removed
EL_HOME <- EL_HOME[-which(EL_HOME$Boligtilstand == "Daarlig"),]
EL_HOME$Boligtilstand <- factor(EL_HOME$Boligtilstand)
```

\newpage
```{r "residuals and leverage7", warning=FALSE}
summary(EL_HOME$AntalToiletter)
```
\newpage

```{r "residuals and leverage8", warning=FALSE}
EL_HOME$AntalToiletter <- factor(EL_HOME$AntalToiletter)
```

# Reducing parameters


```{r "reducing parameters1", warning=FALSE}
mod1 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Opfoerelsesaar + Liggetid + Salgsaar + Season + 
             Fracentrum,
           data = EL_HOME)
```

\newpage
```{r "reducing parameters2", warning=FALSE}
summary(mod1)

AIC(mod1)
```

\newpage
```{r "reducing parameters3", warning=FALSE}
# attempt at removal of insignificant parameters:
# Season, Liggetid, Antal toiletter
mod2 <- lm(log(Kontantpris) ~ Boligtilstand + Boligareal + Altan +
             OmbygningSket + Opfoerelsesaar + Salgsaar + Fracentrum,
           data = EL_HOME)
```

\newpage
```{r "reducing parameters4.1", warning=FALSE}
summary(mod2)

AIC(mod2)
```
\newpage

```{r "reducing parameters4.2", warning=FALSE}
anova(mod1,mod2)
# the p-value is above 0.05 so the parameters can safely be removed
```
\newpage

```{r "reducing parameters5", warning=FALSE}
# attempt at removal of insignificant parameters:
# OmbygningSket
mod3 <- lm(log(Kontantpris) ~ Boligareal + Altan + Opfoerelsesaar + Boligtilstand +
             Salgsaar + Fracentrum, data = EL_HOME)
```
\newpage

```{r "reducing parameters6.1", warning=FALSE}
summary(mod3)

AIC(mod3)

sqrt(mean((exp(predict(mod3)) - EL_HOME$Kontantpris)^2))

#RMSE
sqrt(mean(residuals(mod3)^2))
```
\newpage

```{r "reducing parameters6.2", warning=FALSE}
anova(mod2,mod3)
# the p-value says that OmbygningSket can safely be removed
```
\newpage

```{r "reducing parameters7", warning=FALSE}
# attempt at removal of insignificant parameters:
# Opfoerelsesaar
mod4 <- lm(log(Kontantpris) ~ Boligareal + Altan + Boligtilstand +
             Salgsaar + Fracentrum, data = EL_HOME)
```
\newpage

```{r "reducing parameters8.1", warning=FALSE}
summary(mod4)

AIC(mod4)
```
\newpage

```{r "reducing parameters8.2", warning=FALSE}
anova(mod3,mod4)
# the p-value says that Opfoerelsesaar can not be removed
```
\newpage

# Check mod3

```{r "residual and leverage check of mod31"}
res <- rstudent(mod3)
pred <- predict(mod3)
index <- seq_along(EL_HOME$Kontantpris)
```
\newpage

```{r}
ggplot(cbind(EL_HOME, res, pred, index), aes(index, res)) +
  geom_point() +
  theme_bw()
```
\newpage


```{r "residual and leverage check of mod32"}
ggplot(cbind(EL_HOME, res, pred), aes(pred, res)) +
  geom_point() +
  theme_bw()
```
\newpage

```{r "residual and leverage check of mod33"}
hat <- data.frame(hatvalues(mod3))

ggplot(hat, aes(hat$hatvalues.mod3.)) +
  geom_histogram() +
  theme_bw()

```
\newpage

# Cross validation K fold

```{r "cross validation k fold1"}
train.control <- trainControl(method = "cv", 
                              number = 5)
# Train the model
model <- train(log(Kontantpris) ~ Boligareal + Altan + Boligtilstand +
                 Salgsaar + Opfoerelsesaar + Fracentrum, 
               data = EL_HOME, method = "lm",
               trControl = train.control)
# Summarize the results
# print(model)

model

```

\newpage
```{r "cross validation k fold2"}
AIC(model$finalModel)
mod3$coefficients
model$finalModel
model$results

```
\newpage
# Cross validation leave one out

```{r "cross validation leave one out1"}
# Define training control
train.control <- trainControl(method = "LOOCV")
# Train the model
model_loocv <- train(log(Kontantpris) ~ Boligareal + Altan + Boligtilstand +
                       Salgsaar + Opfoerelsesaar + Fracentrum,
                     data = EL_HOME, method = "lm",
                     trControl = train.control)
# Summarize the results
# print(model_loocv)


model_loocv
```

\newpage
```{r "cross validation leave one out2"}
AIC(model_loocv$finalModel)
mod3$coefficients
model_loocv$finalModel
model_loocv$results

```
\newpage

```{r}
anova(mod1,mod3)
```

```{r eval=FALSE, include=FALSE}

mod0 <- lm(log(Kontantpris) ~ Boligareal, data = EL_HOME)
AIC(mod0)
AIC(mod4)
AIC(mod3)
AIC(mod2)
AIC(mod1)

```



```{r eval=FALSE, include=FALSE}

centrum <- c(9.921747, 57.048820)
data_m_centrum <- data.frame(long = HOME$long - centrum[1], lat = HOME$lat
                             - centrum[2])
length_centrum <- apply(data_m_centrum, 1, norm, type = "2")
HOME$Fracentrum <- length_centrum

HOME_center_aab <- HOME %>% filter(length_centrum < 0.2)
pal <- colorQuantile("YlOrRd", NULL, n = 5)
leaflet(HOME_center_aab) %>% 
  addTiles() %>%
  addCircleMarkers(color = ~pal(Kontantpris), radius = 1)


```


```{r eval=FALSE, include=FALSE}

X <- model.matrix(mod5)
H <- X %*% solve(t(X) %*% X) %*% t(X)
H_anden <- H %*% H
norm(H - H_anden, "F")
H_minus <- t(H) - H

```

```{r}
pris_low <- EL_HOME$Kontantpris * 0.8
pris_high <- EL_HOME$Kontantpris * 1.2
sum(exp(pred) < pris_high & exp(pred) > pris_low) / length(EL_HOME$Kontantpris)

```



# Forward method


```{r include=FALSE}
mod0 <- lm(log(Kontantpris) ~ Boligareal, data = EL_HOME)
summary(mod0)
mod1 <- lm(log(Kontantpris) ~ Boligareal + Salgsaar, data = EL_HOME)
summary(mod1)
mod2 <- lm(log(Kontantpris) ~ Boligareal + Salgsaar + Fracentrum, data = EL_HOME)
summary(mod2)
mod3 <- lm(log(Kontantpris) ~ Boligareal + Salgsaar + Fracentrum + 
             Alder, data = EL_HOME)
summary(mod3)
mod4 <- lm(log(Kontantpris) ~ Boligareal + Salgsaar + Fracentrum + 
             Alder + OmbygningSket, data = EL_HOME)
summary(mod4)
mod5 <- lm(log(Kontantpris) ~ Boligareal + Salgsaar + Fracentrum + 
             Alder + OmbygningSket + Boligtilstand, data = EL_HOME)
summary(mod5)

mod6 <- lm(log(Kontantpris) ~ Boligareal + Salgsaar + Fracentrum + 
             Alder + OmbygningSket + Boligtilstand + Altan, data = EL_HOME)
```

\newpage
```{r}
summary(mod6)

AIC(mod6)

sqrt(mean((exp(predict(mod6)) - EL_HOME$Kontantpris)^2))

#RMSE
sqrt(mean(residuals(mod6)^2))
```


\newpage


```{r}
trainIndex <- createDataPartition(EL_HOME$Kontantpris, p = .7, 
                                  list = FALSE, 
                                  times = 1)
EL_HOME_TRAIN <- EL_HOME[trainIndex,]

```



```{r}
train.control <- trainControl(method = "cv", 
                              number = 5)
# Train the model
model <- train(log(Kontantpris) ~ Boligareal + Salgsaar + Fracentrum + 
             Alder + OmbygningSket + Boligtilstand + Altan, 
               data = EL_HOME, method = "lm",
               trControl = train.control)

model
```

\newpage

```{r}
AIC(model$finalModel)
mod6$coefficients
model$finalModel
model$results
```



\newpage

```{r}
train.control <- trainControl(method = "LOOCV")
# Train the model
model_loocv <- train(log(Kontantpris) ~ Boligareal + Salgsaar + Fracentrum + 
             Alder + OmbygningSket + Boligtilstand + Altan,
                     data = EL_HOME, method = "lm",
                     trControl = train.control)

model_loocv
```
\newpage

```{r}
AIC(model_loocv$finalModel)
mod6$coefficients
model_loocv$finalModel
model_loocv$results
```
\newpage

```{r}
# mod3 <- lm(log(Kontantpris) ~ Boligareal + Altan + Opfoerelsesaar + Boligtilstand +
#              Salgsaar + Fracentrum, data = EL_HOME)
# 
# mean(exp(predict(mod3)) - EL_HOME$Kontantpris)
# mean(exp(predict(mod6)) - EL_HOME$Kontantpris)
# 
# mean(exp(predict(model, EL_HOME[-trainIndex,])) - EL_HOME[-trainIndex,]$Kontantpris)
# mean(exp(predict(model_loocv, EL_HOME[-trainIndex,])) -
#        EL_HOME$Kontantpris[-trainIndex])
# mean(exp(predict(model, EL_HOME[trainIndex,])) - EL_HOME[trainIndex,]$Kontantpris)
# mean(exp(predict(model_loocv, EL_HOME[trainIndex,])) -
#        EL_HOME$Kontantpris[trainIndex])
# 
# summary(mod3)["adj.r.squared"]
# summary(mod6)["adj.r.squared"]


```


```{r}

res <- rstudent(mod6)
pred <- predict(mod6)
index <- seq_along(EL_HOME$Kontantpris)
```
\newpage

```{r}
pris_low <- EL_HOME$Kontantpris * 0.8
pris_high <- EL_HOME$Kontantpris * 1.2
sum(exp(pred) < pris_high & exp(pred) > pris_low) / length(EL_HOME$Kontantpris)
```

\newpage

```{r}
ggplot(cbind(EL_HOME, res, pred, index), aes(index, res)) +
  geom_point() +
  theme_bw()
```
\newpage


```{r "residual and leverage check of mod321"}
ggplot(cbind(EL_HOME, res, pred), aes(pred, res)) +
  geom_point() +
  theme_bw()


```



```{r}

a <- ggplot(cbind(EL_HOME, res, pred), aes(Boligareal, res)) +
  geom_point() +
  theme_bw(base_size = 15)

```

```{r}

b <- ggplot(cbind(EL_HOME, res, pred), aes(Salgsaar, res)) +
  geom_point() +
  theme_bw(base_size = 15)


```


```{r}

c <- ggplot(cbind(EL_HOME, res, pred), aes(Fracentrum, res)) +
  geom_point() +
  theme_bw(base_size = 15)


```

```{r}

d <- ggplot(cbind(EL_HOME, res, pred), aes(Alder, res)) +
  geom_point() +
  theme_bw(base_size = 15)

```


```{r}

e <- ggplot(cbind(EL_HOME, res, pred), aes(OmbygningSket, res)) +
  geom_point() +
  theme_bw(base_size = 15)

```

```{r}

f <- ggplot(cbind(EL_HOME, res, pred), aes(Boligtilstand, res)) +
  geom_point() +
  theme_bw(base_size = 15)

```

```{r}
g <- ggplot(cbind(EL_HOME, res, pred), aes(Altan, res)) +
  geom_point() +
  theme_bw(base_size = 15)
```

```{r "multiplot function", include=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```


```{r}

layout <- matrix(c(1,2,5,3,4,8,6,7),4,2,byrow=TRUE)
multiplot(g, b, c, d, e, f, a, cols = 2)
```

```{r}

anova(mod6)
summary(mod6)

```

```{r}

anova(mod1)

mod2 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Opfoerelsesaar + Salgsaar + Season + 
             Fracentrum,
           data = EL_HOME)

anova(mod2)

mod3 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Opfoerelsesaar + Salgsaar + 
             Fracentrum,
           data = EL_HOME)
anova(mod3)
summary()

mod4 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Salgsaar + Fracentrum + Alder,
           data = EL_HOME)
anova(mod4)

summary(mod4)

train.control <- trainControl(method = "LOOCV")
# Train the model
model_loocv <- train(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter +
                       Altan + OmbygningSket + Salgsaar + Fracentrum + Alder,
                     data = EL_HOME, method = "lm",
                     trControl = train.control)

model_loocv

```

```{r}
EL_HOME[1,]
nyt <- EL_HOME[1,]
nyt$Boligareal <- nyt$Boligareal +1
coefficients(mod6)

exp(predict(mod6, nyt)) - exp(predict(mod6, EL_HOME[1,]))

nyt2 <- EL_HOME[2,]; nyt2$Boligareal <- nyt2$Boligareal +1
exp(predict(mod6, nyt2)) - exp(predict(mod6, EL_HOME[2,]))

```