---
title: "Master"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
library(GGally)
library(caret) # for cross validation
library(readr)
library(dataMaid)
library(corrplot)
```

```{r "data import", include=FALSE}
load("C:/Users/rasmu/OneDrive - Aalborg Universitet/0P5/R-projekt-p5/HOME dataset updated/HOME_med_koordinater.Rda")
#Tænker at man kan nøjes med at inkludere kode og plots ved brug af \includepdf i overleaf, altså ved at slette den smule lim tekst jeg skriver mellem code chunks og så knitte til pdf. Eller bare kopiere relevante chunks ind i andet dokument for at knitte.
```

Firstly dataMaid is used to get a generel feel for the data, this will not be included in the project as it is too large.

```{r "data report dataMaid", eval=FALSE, include=FALSE}
# makeDataReport(HOME)
```

From the data report a number of issues are noticed with the classes of the parameters, such as parameters which should be integers and so on. Aswell as parameters with a lot of missing values.

```{r "fixing parameter classes"}
# select only "ejerlejligheder" in "Aalborg" and
# remove parameters with too many NA's, and removing rows with NA's

EL_HOME <- HOME %>% filter(EjdType == "Ejerlejlighed") %>% 
  filter(Postnr == 9000) %>% 
  select(-c(NytKokken, NyeVinduer, NytBadevaerelse, 
            AntalSovevaerelser, TaetVedVand)) %>% na.omit()


EL_HOME$OmbygningSket <- factor(EL_HOME$OmbygningSket)
EL_HOME$Hoejhus <- factor(EL_HOME$Hoejhus)
EL_HOME$Storgrund <- factor(EL_HOME$Storgrund)
EL_HOME$Kaelder <- factor(EL_HOME$Kaelder)
EL_HOME$MangeToil <- factor(EL_HOME$MangeToil)
EL_HOME$Salgsaar <- as.integer(EL_HOME$Salgsaar)
EL_HOME$GarageCarport <- factor(EL_HOME$GarageCarport)
EL_HOME$Altan <- factor(EL_HOME$Altan)
EL_HOME$AntalToiletter <- factor(EL_HOME$AntalToiletter)
EL_HOME$AntalPlan <- factor(EL_HOME$AntalPlan)
EL_HOME$Boligtilstand <- factor(EL_HOME$Boligtilstand)
EL_HOME$EjdType <- factor(EL_HOME$EjdType)
EL_HOME$Liggetid <- as.integer(EL_HOME$Liggetid)


```

Now the distribution of "Kontant pris" is checked, which should be normal for a GLM.

```{r "check normal distribution"}

qqnorm(EL_HOME$Kontantpris)
qqline(EL_HOME$Kontantpris)

# A logarithmic transformation is tried
qqnorm(log(EL_HOME$Kontantpris))
qqline(log(EL_HOME$Kontantpris))

ggplot(EL_HOME, aes(log(Kontantpris))) +
  geom_density(size = 1.2) +
  theme_bw()

ggplot(EL_HOME, aes((Kontantpris))) +
  geom_density(size = 1.2) +
  theme_bw()

```

# Parameter for season.

Also removing parameters irrelevant for condoes

```{r "parameter for season"}
# Gathering quarter parameters to one
EL_HOME$Season <- EL_HOME$Kvartal1 + EL_HOME$Kvartal2 * 2 + EL_HOME$Kvartal3 * 3 + 
  EL_HOME$Kvartal4 *4
EL_HOME$Season <- factor(EL_HOME$Season)

# removing Kvartal parameters, and also other irrelevant parameters for our subset
# of the data
EL_HOME <- EL_HOME %>% select(-c(Kvartal1, Kvartal2, Kvartal3, Kvartal4,
                                 GarageCarport, x, y, long, lat, Hoejhus,
                                 Storgrund, Kaelder, Kaelderareal, Grundareal,
                                 AntalPlan))

```

# Generel Linear model

Firstly a model with all relevant parameters is tried. For example Alder is removed because it isn't linearly independent of Opfoerelsesaar and Salgsaar.

```{r "initial model"}
mod1 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Opfoerelsesaar + Liggetid + Salgsaar + Season,
           data = EL_HOME)
summary(mod1)
```

We will now check for outliers as a result of mistakes in the data.


```{r "residuals and leverage", warning=FALSE}

res <- rstudent(mod1)
pred <- predict(mod1)

ggplot(cbind(EL_HOME, res, pred), aes(Kontantpris, res)) +
  geom_point() +
  theme_bw()

ggplot(cbind(EL_HOME, res, pred), aes(pred, res)) +
  geom_point() +
  theme_bw()

res_large <- which(abs(res) > 4.5)
EL_HOME[res_large,]

hat <- data.frame(hatvalues(mod1))

ggplot(hat, aes(hat$hatvalues.mod1.)) +
  geom_histogram() +
  theme_bw()

hat_large <- which(hat > 0.2)
EL_HOME[hat_large,]
#these condoes seem to be mistakes so are removed
EL_HOME <- EL_HOME[-res_large,]
EL_HOME <- EL_HOME[-hat_large,]
# the model is then reduced

# Some factor parameters might have changed
summary(EL_HOME$Boligtilstand)
# Factor with only 1 entry is removed
EL_HOME <- EL_HOME[-which(EL_HOME$Boligtilstand == "Daarlig"),]
EL_HOME$Boligtilstand <- factor(EL_HOME$Boligtilstand)
summary(EL_HOME$AntalToiletter)
EL_HOME$AntalToiletter <- factor(EL_HOME$AntalToiletter)

```

# Reducing parameters


```{r "reducing parameters", warning=FALSE}
mod1 <- lm(log(Kontantpris) ~ Boligtilstand +Boligareal + AntalToiletter + Altan +
             OmbygningSket + Opfoerelsesaar + Liggetid + Salgsaar + Season,
             data = EL_HOME)
summary(mod1)

# attempt at removal of insignificant parameters:
# Season, Liggetid, Antal toiletter
mod2 <- lm(log(Kontantpris) ~ Boligtilstand + Boligareal + Altan +
             OmbygningSket + Opfoerelsesaar + Salgsaar, data = EL_HOME)
summary(mod2)
anova(mod1,mod2)
# the p-value is above 0.05 so the parameters can safely be removed

# attempt at removal of insignificant parameters:
# OmbygningSket
mod3 <- lm(log(Kontantpris) ~ Boligareal + Altan + Opfoerelsesaar + Boligtilstand +
             Salgsaar, data = EL_HOME)
summary(mod3)
anova(mod2,mod3)
# the p-value says that OmbygningSket can safely be removed

# attempt at removal of insignificant parameters:
# Opfoerelsesaar
mod4 <- lm(log(Kontantpris) ~ Boligareal + Altan + Boligtilstand +
             Salgsaar, data = EL_HOME)
summary(mod4)
anova(mod3,mod4)
# the p-value says that Opfoerelsesaar can safely be removed
# now only Salgsaar is a parameter so Alder can be added without 
# causing linear dependency

mod5 <- lm(log(Kontantpris) ~ Boligareal + Altan + Boligtilstand +
             Salgsaar + Alder, data = EL_HOME)
summary(mod5)
anova(mod4,mod5)
# the p-value shows that Alder can not be removed

# the model can not be reduced to less than mod5

```


# Check mod5

```{r "residual and leverage check of mod5"}
res <- rstudent(mod5)
pred <- predict(mod5)

ggplot(cbind(EL_HOME, res, pred), aes(log(Kontantpris), res)) +
  geom_point() +
  theme_bw()

ggplot(cbind(EL_HOME, res, pred), aes(pred, res)) +
  geom_point() +
  theme_bw()

hat <- data.frame(hatvalues(mod5))

ggplot(hat, aes(hat$hatvalues.mod5.)) +
  geom_histogram() +
  theme_bw()


```


# Cross validation K fold

```{r "cross validation k fold"}

train.control <- trainControl(method = "repeatedcv", 
                              number = 5, repeats = 10)
# Train the model
model <- train(log(Kontantpris) ~ Boligareal + Altan + Boligtilstand +
             Salgsaar + Alder, 
               data = EL_HOME, method = "lm",
               trControl = train.control)
# Summarize the results
# print(model)


model$finalModel
mod5$coefficients
model$results

```

# Cross validation leave one out

```{r "cross validation leave one out"}

# Define training control
train.control <- trainControl(method = "LOOCV")
# Train the model
model_loocv <- train(log(Kontantpris) ~ Boligareal + Altan + Boligtilstand +
             Salgsaar + Alder,
                     data = EL_HOME, method = "lm",
                     trControl = train.control)
# Summarize the results
# print(model_loocv)

model_loocv$finalModel
mod5$coefficients
model_loocv$results

```


```{r}
ordered_data <- arrange(EL_HOME, Kontantpris)

yc <- as.data.frame(apply(predict(mod5, ordered_data, interval = "confidence"), 2, exp))
names(yc) <- c("pred", "lower_conf", "upper_conf")
ypi <- as.data.frame(apply(predict(mod5, ordered_data, interval = "prediction"),2,exp)) %>% 
  select(-fit)
names(ypi) <- c("lower_pred", "upper_pred")

plot_data <- bind_cols(ordered_data, yc, ypi, index = seq_along(EL_HOME$Kontantpris))

subset <- seq(1, length(EL_HOME$Kontantpris), length.out = 100)
plot_data[subset,]

ggplot(plot_data[,], aes(index, Kontantpris)) +
  geom_point() +
  # geom_point(aes(index, pred), color = "blue") +
  # geom_point(aes(index, lower_conf), color = "red") +
  # geom_point(aes(index, upper_conf), color = "red") +
  # geom_point(aes(index, lower_pred), color = "green") +
  # geom_point(aes(index, upper_pred), color = "green") +
  geom_smooth(aes(index, lower_conf)) +
  geom_smooth(aes(index, upper_conf)) +
  geom_smooth(aes(index, lower_pred)) +
  geom_smooth(aes(index, upper_pred)) +
  geom_smooth(aes(index, pred))
  

```






